---
title: "p8105_hw2_cq2207"
author: "Carolina Q Cardoso"
date: "10/2/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r problem1}


library(tidyverse)
library(readxl)

#Mr. Trash Wheel

mr_trash = read_excel(path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1, col_names=TRUE) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls =  as.integer(round(sports_balls, digits=0))) %>%
  select(-x15, -x16, -x17)

#Precipitation 2017

p_2017 = read_excel(path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 6, col_names=TRUE, skip=1) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year=2017)
  
#Precipitation 2018

p_2018 = read_excel(path = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 5, col_names=TRUE, skip=1) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year=2018)

#combine 2017 and 2018 precipitation data 

p_combined = full_join(p_2017,p_2018, by = "month") %>%
  mutate(month = month.name) %>%
  rename(`year_2017` = year.x, `year_2018` = year.y, total_2017 = total.x, total_2018 = total.y) 
  
```

#write paragraph about data#

The `mr_trash` dataset contains information on waste collected in dumpsters. It contains `r nrow(mr_trash)` observations, and includes information such as the total weight (`weight_tons`) and volume (`volume_cubic_yards`) collected by each individual dumpster, as well as number of specific items discarded in the dumpsters, such as `glass_bottles`, `grocery_bags`, and `chip_bags`. From the dataset we can see that the median number of sports balls in a dumpster in 2017 was `r median(pull(mr_trash,sports_balls))`.
The `p_combined` dataset reports monthly precipitation for the years of 2017 and 2018, and contains `r nrow(p_combined)` observations. It includes the `month`, `year`, and `total` precipitation for that month in a given year, in inches. 
The total precipitation in 2018 was `r sum(pull(p_combined,total_2018))`. 

```{r problem2}

#pols-month.csv data

pols = 
    read_csv("./data/pols-month.csv") %>%
    janitor::clean_names() %>%
    separate(mon, into = c("year", "month", "day"), sep = "-") %>%
    mutate(month = month.name[as.numeric(month)]) %>%
    mutate(prez_all = if_else(prez_gop=="1", "gop", "dem")) %>%
    select(-prez_gop, -prez_dem,-day)

head(pols)

#snp.csv

snp = 
    read_csv("./data/snp.csv") %>%
    janitor::clean_names() %>%
    separate(date, into = c("month", "day", "year"), sep = "/") %>%
    mutate(month = month.name[as.numeric(month)]) %>%
    select(-day) %>%
    select(year, month, close)

head(snp)

#unemployment.csv

unemployment_wide = 
    read_csv("./data/unemployment.csv") %>%
    janitor::clean_names() 
unemp <- gather(unemployment_wide, month, unemp_rate, jan:dec, factor_key=TRUE) %>%
    mutate(month = month.name[as.numeric(month)]) %>%
    mutate(year = as.numeric(year))

#merge 3 data sets

first_join = full_join(pols, snp, by=c("year","month")) %>%
  mutate(year = as.numeric(year))
join_complete = full_join(first_join, unemp, by=c("year","month"))

```

#write about resulting dataset

The dataset `join_complete` contains `r nrow(join_complete)` observations and `r ncol(join_complete)` variables. The key variables in this data set are `year` and `month`.
It combines three datasets. The first is `pols`, which describes the number of political servants (governors, senators, and congresspeople) belonging to either the republican or democratic party; as well as the political affiliation of the president, in a given year and month. The data set `snp`, describes the closing value of the SNP on the first weekday of that month and year. And, finally, the `unemployment` dataset describes the unemployment rate in a given year and month.
The range of years in the dataset go from `r min(pull(join_complete,year))` to `r max(pull(join_complete,year))`. 

```{r problem3}

#clean dataset

baby_names = 
    read_csv("./data/Popular_Baby_Names.csv") %>%
    janitor::clean_names() %>%
    mutate_all(tolower) %>%   
    mutate(
    ethnicity = replace(ethnicity, ethnicity == "asian and paci", "asian and pacific islander"),
    ethnicity = replace(ethnicity, ethnicity == "black non hisp", "black non hispanic"),
    ethnicity = replace(ethnicity, ethnicity == "white non hisp", "white non hispanic")) %>% 
  distinct() 


#make Olivia table

olivia_rank = 
filter(baby_names, childs_first_name == "olivia") %>% 
  select(-gender, -childs_first_name, -count) 

library(knitr)
olivia_table <- spread(olivia_rank,year_of_birth,rank) %>% 
knitr::kable() %>% 
    print()

#make similar table for popular male names

male_popular = 
filter(baby_names, gender == "male", rank == '1') %>% 
  select(-gender, -rank, -count) 

male_popular_table <- spread(male_popular,year_of_birth,childs_first_name) %>% 
knitr::kable() %>% 
    print()


```